### 2022-04-10 paper_5.rmd
init commit

---
title: "W(ha/il)t's the (wo/ba)rd?" # TODO
subtitle: "A cross-sectional analysis of English word pairs that satisfy the properties of the Split Decisions word puzzle"
author: "Oliver Daniel"
abstract: >
  From three English wordlist corpora, pairs of words are generated which might appear in the Split Decisions word puzzle as published by the New York Times, i.e., at least 5 characters in length and differing by exactly two consecutive characters. The corpora are contrasted among themselves, and then the lexical and phonological properties of those words which form pairs – and those which do not – are analyzed for causative insights. Such insights will be alluded to here once discovered.

date: "April 6, 2022"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: false
knit: (function(input, ...) {
    rmarkdown::render(
    input,
    output_file = "out/paper_5.pdf",
    envir = globalenv(),
    ...
    )
  })
bibliography: in/references.bib
link-citations: true
---
\\par \\textbf{Keywords}: crosswords, split decisions, computational linguistics

Code available on [GitHub (link)](https://github.com/oliver-daniel/inf_313_final_paper).
```{r setup, echo=FALSE}
knitr::opts_chunk$set(
    echo = FALSE, warning = FALSE,
    out.width = "100%", out.align = "center", dir
)
options(knitr.graphics.error = FALSE)
```

```{r libraries, include=FALSE}
if (!require(tidyverse)) {
    install.packages("tidyverse")
}
if (!require(kableExtra)) {
    install.packages("kableExtra")
}

library(tidyverse)
library(knitr) # TODO CITE
library(kableExtra) # TODO CITE
```

## Introduction
A few times a year, the New York Times games department releases a new issue of Split Decisions, a crossword-like word puzzle. Similar to crosswords, the game board consists of boxes arranged in sequence, one for each letter of a word running either left-to-right or top-to-bottom, intersecting with other words at right angles. However, instead of cryptic descriptions of correct answers, Split Decisions solvers are only given pairs of two-letter substrings for each word, either of which serves to complete a valid English word. For example, given the following diagram:

<!-- TODO: visual example -->
$${PL \choose GR} \_\ \_\ \_\ \_$$

A correct fill for the remaining boxes might be __EASE, creating both PLEASE and GREASE. For the remainder of this paper, such a pair of words will be referred to as a *split pair*. Conversely, an incorrect fill might be __IERS, which produces the valid word PLIERS and the invalid word \*GRIERS[^1]. Note that each combination of clues (hereon a *split*) might accept several different fills for the common substring, e.g., __UNGE (PLUNGE, GRUNGE); however, as with crosswords, only one solution will also satisfy all the crossing constraints produced by the rest of the puzzle.

Note that, although a letter may appear on both sides of a split, say, (IT/TA), the letter will *never* appear at the same index in both words, e.g., (IT/AT), as such a pair of words would only differ by a single letter instead of two. Also, although such a puzzle would be fascinating to both solve and analyze, crossing words do *not* go through split pairs, only common letters.

Upon understanding the rules of the puzzle, one may start to wonder: what sorts of words form split pairs? Are some splits more common than others? How do the properties of English, as both a spoken and written language, affect the guessability of a given split pair? In this paper, we perform preliminary statistical analysis to investigate these questions, using a variety of computational methods.

## Data
I used [@R], [@tidyverse], and [@ggplot2].
The wordlists come from GNU/Linux (specifically `/usr/share/dict/words` on an Ubuntu distribution), [@collins], and [@nwl2020].

```{r read-data}
corpus.linux <- read.table("in/corpus/linux.txt")
corpus.collins <- read.table("in/corpus/collins.txt")
corpus.nwl <- read.table("in/corpus/NWL2020.txt")

dat.linux <-
    read.csv("in/cache/linux.csv") |>
    tibble() |>
    mutate(
        corpus = "linux",
    )

dat.collins <-
    read.csv("in/cache/collins.csv") |>
    tibble() |>
    mutate(
        corpus = "collins",
        start = start + 1, # Made this column zero-indexed by accident :)
        end = end + 1
    )

dat.nwl <-
    read.csv("in/cache/NWL2020.csv") |>
    tibble() |>
    mutate(corpus = "nwl")

dat.all <- function() {
    rbind(
        dat.linux,
        dat.collins,
        dat.nwl
    )
}
```

```{r}
most_common <- function(col, n = 1) {
    table(col) |>
        sort() |>
        names() |>
        tail(n)
}

PATTERN <- "[a-z]{5,}"

corpus_stats <- rbind(
    tibble(
        corpus = "linux",
        all_words = corpus.linux |> nrow(),
        valid_words = corpus.linux |>
            filter(str_detect(V1, PATTERN)) |>
            nrow(),
        pairs_count = dat.linux |> nrow()
    ),
    tibble(
        corpus = "collins",
        all_words = corpus.collins |> nrow(),
        valid_words = corpus.collins |>
            filter(str_detect(V1, PATTERN)) |>
            nrow(),
        pairs_count = dat.collins |> nrow()
    ),
    tibble(
        corpus = "nwl",
        all_words = corpus.nwl |> nrow(),
        valid_words = corpus.nwl |>
            filter(str_detect(V1, PATTERN)) |>
            nrow(),
        pairs_count = dat.nwl |> nrow()
    )
)

split_pair_counts <- dat.all() |>
    mutate(split_pair = paste0(split_x, "/", split_y)) |>
    group_by(corpus) |>
    count(split_pair, name = "split_pair_count") |>
    slice_max(n = 1, order_by = split_pair_count, with_ties = F)

split_x_counts <- dat.all() |>
    group_by(corpus) |>
    count(split_x, name = "split_x_count") |>
    slice_max(n = 1, order_by = split_x_count, with_ties = F)

split_y_counts <- dat.all() |>
    group_by(corpus) |>
    count(split_y, name = "split_y_count") |>
    slice_max(n = 1, order_by = split_y_count, with_ties = F)
```

```{r}
corpus_stats |>
    left_join(split_x_counts, by = "corpus") |>
    left_join(split_y_counts, by = "corpus") |>
    left_join(split_pair_counts, by = "corpus") |>
    mutate(
        split_x = paste0(split_x, " (", split_x_count, ")"),
        split_y = paste0(split_y, " (", split_y_count, ")"),
        split_pair = paste0(split_pair, " (", split_pair_count, ")")
    ) |>
    select(corpus, all_words, valid_words, pairs_count, split_x, split_y, split_pair) |>
    kable(
        booktabs = TRUE,
        col.names = c("Corpus", "Total",
            "Valid", "Total", "First split",
            "Second split", "Split pair")
    ) |>
    kable_classic() |>
    add_header_above(c(" " = 1, "Words" = 2, "Pairs" = 1, "Most frequent (occurrences)" = 3))
```

## Results

## Discussion

\newpage
## Appendix
<!-- TODO: python script -->

\newpage
## References

<!-- Footnotes -->
[^1]: The asterisk prefix will be used throughout this paper to indicate incorrect/invalid words.