### 2022-04-10 scripts/exploration.rmd

```{r}
library(tidyverse)
library(stringdist)
```

```{r}
dat.linux <-
  read.csv("in/cache/linux.csv") |>
  tibble() |>
  add_column(corpus = "linux")

dat.collins <-
  read.csv("in/cache/collins.csv") |>
  tibble() |>
  add_column(corpus = "collins") |>
  mutate(
    start = start + 1, # Made this column zero-indexed by accident :)
    end = end + 1
  )

dat.nwl <-
  read.csv("in/cache/NWL2020.csv") |>
  tibble() |>
  add_column(corpus = "nwl")
# Defer binding + memory allocation until last possible second --
# no clue if that actually does anything
dat.all <- function() {
  rbind(
    dat.linux,
    dat.collins,
    dat.nwl
  )
}
```

```{r}
dat.all() |>
  # slice_sample(prop = 1) |>
  # group_by(corpus) |>
  # add_count(length) |>
  ggplot(aes(x = length, fill = corpus)) +
  geom_bar() +
  scale_x_continuous(breaks = 5:21) +
  scale_y_continuous(trans = "log10") +
  theme(panel.grid.minor.x = element_blank())
```

```{r}
# most common pairings, in original
# lexical order
dat.all() |>
  filter(end == length) |>
  select(x, y, split_x, split_y, length, end) |>
  count(split_x, split_y) |>
  arrange(desc(n))
```

```{r}
suffixes <- dat.all() |>
  filter(end == length)

suffix_counts <- c(
  suffixes$split_x,
  suffixes$split_y
) |>
  table() |>
  data.frame() |>
  tibble() |>
  arrange(desc(Freq)) |>
  rowid_to_column(var = "rank")

## Interesting:
# the top 6 suffixes are all
# declensions: es, ed, er, (i)ng, (i)on, ly
suffix_counts |>
  ggplot(aes(x = log10(rank), y = log10(Freq))) +
  geom_point() +
  scale_x_continuous(limits = c(0, 5)) +
  scale_y_continuous(limits = c(0, 5)) +
  coord_fixed()
```

```{r}
# let's take a look at that more generally

splits_by_type <- dat.all() |>
  mutate(
    type = case_when(
      start == 1 ~ "prefix",
      end == length ~ "suffix",
      TRUE ~ "middle"
    )
  )

rbind(
  splits_by_type |> select(split = split_x, type),
  splits_by_type |> select(split = split_y, type)
) |>
  count(split, type) |>
  group_by(type) |>
  mutate(
    rank = rank(-n, ties.method = "first"),
    f = max(n) / rank
  ) |>
  ggplot(aes(x = rank, y = n, color = type, )) +
  geom_point() +
  geom_line(aes(x = rank, y = f), linetype = 2) +
  # geom_smooth(method = "lm", se = F, alpha = .5) +
  scale_x_continuous(limits = c(1, 26^2)) +
  scale_y_log10() #+
```

```{r}
# what about /all/ splits?

if (!file.exists("in/cache/corpus_splits.csv")) {
  read.table("in/corpus/linux.txt") |>
    tibble() |>
    mutate(corpus = "linux") |>
    rbind(
      read.table("in/corpus/collins.txt") |>
        tibble() |>
        mutate(corpus = "collins")
    ) |>
    rbind(
      read.table("in/corpus/NWL2020.txt") |>
        tibble() |>
        mutate(corpus = "nwl")
    ) |>
    filter(str_detect(V1, "^[a-z]{5,}$")) |>
    # slice_sample(prop = .2) |>
    rowwise() |>
    mutate(
      splits = map(
        seq(1, str_length(V1) - 1), ~ paste0(
          substr(V1, ., .),
          substr(V1, . + 1, . + 1)
        )
      ) |> paste0(collapse = ",")
    ) |>
    separate_rows(splits, sep = ",") |>
    write.csv("in/cache/corpus_splits.csv")
}

# Of the 26^2 = 676 possible combinations,
# 614 are accounted for
corpus_splits <- read.csv("in/cache/corpus_splits.csv")

corpus_splits |>
  count(corpus, splits) |>
  group_by(corpus) |>
  mutate(
    rank = rank(-n, ties.method = "first"),
    f = max(n) / rank
  ) |>
  ggplot(aes(x = rank, y = n, color = corpus)) +
  geom_point() +
  geom_line(aes(y = f), linetype = 2) +
  geom_vline(aes(xintercept = 425), linetype = 3) +
  scale_x_continuous(limits = c(1, 26^2)) +
  scale_y_log10()
```